---
title: "Basic Examples"
author: "Matt Johnson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE, 
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  comment = "#>"
)

library(dplyr)
library(jsonlite)

```




## A Minimal Example


```{r min}

library(cytoscape)


nodes <- data.frame(id = c('a','b'))
edges <- data.frame(id = 'ab', source = 'a', target = 'b')

cytoscape(nodes = nodes, edges = edges) 

```


## Styling Edges and Nodes

Nodes and edges can all be styled using CSS.

```{r style1}

ct <- cytoscape(nodes = nodes, edges = edges)

ct %>% node_style('background-color' = '#ff0000')

```


```{r style2}

ct %>% edge_style('line-color' = '#ff0000')

```


## Creating data

Every network chart requires two `data.frames` one for the nodes and another for the edges. Often you will already have a `data.frame` with the edge, or `source`, `target`. In the example below we have a `data.frame` of Comtrade waste plastic exports for December 2017 where the source counrty and the destination country are deliniated. These fields need to be renamed to `source` and `target` for `cytoscape` to use it. For the `nodes` `data.frame` you need to create an `id` field that contains all of the unique values found in the `source` and `target` fields of the `edges` `data.frame`. 


```{r plastics}


df <- cytoscape::comtrade

edges <- df %>%
    dplyr::select(source = reporter,
                  target = partner) %>%
    dplyr::mutate(id = paste(source, '_', target))

nodes <- data.frame(id = unique(c(edges$source, edges$target)), stringsAsFactors = FALSE)
           
cytoscape(nodes = nodes, edges = edges) %>% 
  layout('grid', rows = 4)


```



```{r, echo=FALSE, include=FALSE}
# cytoscape(nodes = nodes, edges = edges) %>%
#   layout('breadthfirst', directed = TRUE)
```





## With Geographic Location

You can use coordinates to fix the location of nodes. But the package is not spatially enabled and does not check or transform the coordinates you pass.

```{r location}


coords <- cytoscape::coords %>%
  rename(x = lon,
         y = lat) %>%
  filter(id != 'World') %>%
  mutate(y = -y) # this will depend on coordinates used

edges <- filter(edges, 
                source %in% coords$id,
                target %in% coords$id)

cytoscape(nodes = coords, edges = edges) %>% 
  layout('preset') %>%
  node_style(width = 3, height = 3, 'font-size' = 8) %>%
  edge_style(width = 1)

```


## Image Node Example

Replicate the `Cytograph.js` [images example](http://js.cytoscape.org/demos/images-breadthfirst-layout/). This example does not currently include the touch effects.

```{r image}

img_nodes <- data.frame(id = c('cat','bird','ladybug','aphid','rose','grasshopper','plant','wheat'),
                        images = c('https://farm2.staticflickr.com/1261/1413379559_412a540d29_b.jpg',
                                   'https://farm8.staticflickr.com/7272/7633179468_3e19e45a0c_b.jpg',
                                   'https://farm4.staticflickr.com/3063/2751740612_af11fb090b_b.jpg',
                                   'https://farm9.staticflickr.com/8316/8003798443_32d01257c8_b.jpg',
                                   'https://farm6.staticflickr.com/5109/5817854163_eaccd688f5_b.jpg',
                                   'https://farm7.staticflickr.com/6098/6224655456_f4c3c98589_b.jpg',
                                   'https://farm1.staticflickr.com/231/524893064_f49a4d1d10_z.jpg',
                                   'https://farm3.staticflickr.com/2660/3715569167_7e978e8319_b.jpg'),
                        stringsAsFactors = FALSE)
img_edges <- data.frame(source = c('cat','bird','bird','grasshopper','grasshopper','ladybug','aphid'),
                        target = c('bird','ladybug','grasshopper','plant','wheat','aphid','rose'),
                        stringsAsFactors = FALSE)

cytoscape(nodes = img_nodes,
          edges = img_edges) %>%
  layout(name = 'breadthfirst',
         directed = TRUE,
         padding = 10) %>%
  node_style('height' = 80,
             'width' = 80,
             'background-fit' = 'cover',
             'border-color' = '#000',
             'border-width' = 3,
             'border-opacity' = 0.5,
             'label' = NULL) %>%
  edge_style('curve-style' = 'bezier',
             'width' = 6,
             'target-arrow-shape' = 'triangle',
             'line-color' = '#ffaaaa',
             'target-arrow-color' = '#ffaaaa') %>%
  node_images()




```


## Node Gradient Background

Since `cytoscape` version 3.3.0 it has been possible to include background gradients in node styles. This is included in this package as of version 0.0.2.

```{r gradient}

cytoscape(nodes = img_nodes,
          edges = img_edges) %>%
  layout(name = 'breadthfirst',
         directed = TRUE,
         padding = 10) %>%
  node_style("content" = "data(id)",
              "text-valign" = "center",
              "text-halign" = "center",
              'height' = 80,
              'width' = 80,
              "background-fill" = "radial-gradient",
              'background-fit' = 'cover',
              "background-gradient-stop-colors" = "cyan magenta yellow", 
              "background-gradient-stop-positions" = "25 75 80"
             ) 
```


## Grouping Nodes

[Compound graphs](https://cytoscape.org/cytoscape.js-cola/demo-compound.html) where nodes are grouped are also possible. Add `parent` and `node_color` columns to the nodes `data.frame`. The final step is to add nodes for each of the groups, note the `bind_rows()` function in the example below. Thanks to [gilhornung](https://github.com/gilhornung) for [this example](https://github.com/mrjoh3/cytoscape/issues/8#issuecomment-475142230). 

```{r groups}

df <- cytoscape::comtrade

gp_nodes <- tibble(id = unique(c(df$reporter, df$partner))) %>% 
  # Define node groups
  mutate(parent = ifelse(id %in% c("Nigeria", "Kenya", "South Africa"),
                         "Africa",
                         "Not Africa")) %>% 
  # Define node colors
  mutate(node_color = ifelse(parent == "Africa",
                             "forestgreen",
                             "darkorange")) %>%
  # Need to add the nodes for the two groups, Africa and "Not Africa"
  bind_rows(data_frame(id = c("Africa", "Not Africa"), 
                       node_color="whitesmoke"))
  
gp_edges <- df %>%
    dplyr::select(source = reporter,
                  target = partner) %>%
    dplyr::mutate(id = paste0(source, '-', target))

cytoscape(nodes = gp_nodes, edges = gp_edges) %>% 
  node_style('background-color' = 'data(node_color)') %>% 
  cola_layout(avoidOverlap = T)

```


## Cytoscape-Cola Layout

Constraint based layouts provided through ([cola.js](http://ialab.it.monash.edu/webcola/)) can be used via the [cytoscape-cola plugin](https://github.com/cytoscape/cytoscape.js-cola). All options available in the plugin [API](https://github.com/cytoscape/cytoscape.js-cola#api) are available but have not yet been tested.


### Basic Cola Layout

```{r cola}

cytoscape(nodes = nodes, edges = edges) %>% 
  cola_layout()


```


### Cola Node Allignment

To use the Cola `alignment` variable you need to pass a javascript object or function. You can do this using `htmltools::JS()`.

```{r cola2}

aln_nodes <- data.frame(id = c('cat','bird','ladybug','grasshopper','wheat'),
                        stringsAsFactors = FALSE)
aln_edges <- data.frame(source = c('cat', 'bird',   'bird',       'grasshopper'),
                        target = c('bird','ladybug','grasshopper','wheat'      ),
                        stringsAsFactors = FALSE)

opts = htmlwidgets::JS('function( node ){ ',
                       'console.log(node._private.data.id);',
                       'if (node._private.data.id === "cat") {return { x: 0} }',
                       'else if (node._private.data.id === "bird") {return { x: 0, y: 0 } }',
                       'else if (node._private.data.id === "ladybug") {return { x: 0} }',
                       'else if (node._private.data.id === "wheat") {return {y: 0 } }',
                       'else if (node._private.data.id === "grasshopper") {return {y: 0 } }',
                       '}')

cytoscape(nodes = aln_nodes, edges = aln_edges) %>% 
  cola_layout(alignment = opts)



```


## Shiny Example

A minimal shiny example can be run from the `cytoscape` package of from `inst/shiny/minimum_shiny`

```r

shiny::runApp(system.file('shiny/minimum_shiny', package = 'cytoscape'))

```

## Pass JSON Object

If you already have a complete `JSON` this can be passed through directly as a character string. The `JSON` character is parsed in `javascript` using `JSON.parse()` so it is good practice to first test your `JSON` is properly formed. On the `R` side you can use `jsonlite::fromJSON()` or the web-service http://json.parser.online.fr/ can sometimes give more meaningfull error messages.

```{r json}

json <- '{
"elements":[
    {
      "data": { "id": "a" } 
    },
    {
      "data": { "id": "b" }
    },
    {
      "data": { "id": "ab", "source": "a", "target": "b" }
    }
  ],

"style": [ 
  {
    "selector": "node",
    "style": {
      "background-color": "#666",
      "label": "data(id)"
    }
  },

  {
    "selector": "edge",
    "style": {
      "width": 3,
      "line-color": "#ccc",
      "target-arrow-color": "#ccc",
      "target-arrow-shape": "triangle"
    }
  }
],

"layout": {
  "name": "grid",
  "rows": 1
}
}'

cytoscape(json = json)

```
